A thermodynamic view of dusty protoplanetary disks I: axisymmetric stability

### 2017-01-19 paper.tex
tiny intro edits

There are several situations where the pressure and density of a fluid
are not in phase. The obvious case is if the fluid is subject to
external heating/cooling. For example, in strongly irradiated
protoplanetary disks the
disk temperature  $T(r)$ is time-independent but varies with radius
$r$ from the star \citep{chiang97}. Since $P\propto \rho T$ for an ideal gas, the pressure of a fluid parcel varies with 
position as well. There is no reason to expect $P$ and $\rho$ to
be in phase. In fact, we will show that this is responsible for 
the vertical shear instability \citep{lin15}.

@@

Another possibility, as annotated in Fig. \ref{pdv_cartoon}, is a dusty gas. The
relevant density here is the total density $\rho = \rhod + \rhog$, 
but the fluid pressure $P$ is due to gas only. If dust were 
perfectly-coupled to the gas, then $P$ and $\rho$ would be in phase, 
and there is no work done. However, for finite dust-gas drag, $\rhod$ and
$\rhog$ are not neccessarily in phase. If this causes $P$ to lag
$\rho$, then the positive work done would lead to growing
oscillations. Indeed, we show that this is true for the streaming
instability.

>>>>

There are several situations where the pressure and density of a fluid
are not in phase. The obvious case is if the fluid is subject to
external heating/cooling. For example, in strongly irradiated
protoplanetary disks the
disk temperature  $T(r)$ is time-independent but varies with the cylindrical radius
$r$ from the star \citep{chiang97}. Since $P\propto \rhog T$ for an ideal gas, the pressure of a gas parcel varies with 
position as well. There is no reason to expect $P$ and $\rhog$ to
be in phase. In fact, we will show that this is a fundamental property of 
the vertical shear instability \citep{lin15}.

@@

Another possibility, as annotated in Fig. \ref{pdv_cartoon}, is a dusty gas. The
relevant density here is the total density $\rho = \rhod + \rhog$, 
but the fluid pressure $P$ is due to gas only. If dust were 
perfectly-coupled to the gas, then $P$ and $\rho$ would be in phase, 
and there is no work done. However, for finite dust-gas drag, $\rhod$ and
$\rhog$ are not neccessarily in phase, because dust particles can drift relative to the gas. 
\emph{If} this causes $P$ to lag 
$\rho$, then the positive work done would lead to growing
oscillations. Indeed, we show that this is true for the streaming
instability.
